Parameter estimation

### 2023-04-22 SFE-article/Article_Model_Parameters.tex
20.04
22.04
21.04

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ are the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
We consider a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ denotes the state variables, $\Theta$ represents the parameters, and $u$ is the control variables. Our focus is on the supercritical extraction process, which involves a partially filled extractor with a fixed bed that operates under constant conditions. We assume the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. To model this two-dimensional case, we use the concept of quasi-one-dimensional flow.
To describe the fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent, we use a distributed-parameter model based on \citet{Reverchon1996}. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, and saturation concentration. We estimate these parameters from four experiments at different temperatures and pressures: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. We obtain the parameter values using the maximum likelihood estimation method on yield data under the normal error assumption.

@@

\begin{keywords}
Supercritical extraction \sep Parameter estimation \sep Mathematical modelling
\end{keywords}

@@

\subsection{Governing equations}
The governing equation for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}.  Quasi-one-dimensional flow is a type of fluid flow that is characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as a irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

@@

Based on governing equations, the small discontinuity (defined as $\delta$) in flow properties, shown on figure \ref{fig: Discontinuity_slow_flow}, can be analysed. The analysis follow the work of \citet{Schreier1982}.

\begin{figure}[!ht]
	\centering
	\resizebox{0.95\columnwidth}{!}{%
		\begin{tikzpicture}[]
			\draw (0,2) -- (9,2);	% Top line
			\draw (0,0) -- (9,0); 	% Bottom line
			\draw (4.5,0) -- (4.5,2); 	% Bottom line
			\node at (4,1.5) {$\rho_f$};
			\node at (5.5,1.5) {$\rho_f+\delta\rho_f$};
			\node at (4,1.0) {$P$};
			\node at (5.5,1.0) {$P+\delta P$};
			\node at (4,0.5) {$T$};
			\node at (5.5,0.5) {$T+\delta T$};
			\node at (1,1.0) {$v \rightarrow$};
			\node at (8,1.0) {$v + \delta v \rightarrow$};
	\end{tikzpicture} }
	\caption{Small discontinuity in one-dimensional flow}
	\label{fig: Discontinuity_slow_flow}
\end{figure}

The discontinuity is presumed to be at rest relative and the balance equations become

@@

These relations are equally valid if the two regions are separated by a regions of finite width rather than a discontinuity.

@@

By combining momentum equation with above equation we get

@@

If the flow is presumed to be isentropic, $dP/d\rho = c^2$, so $v^2=c^2$, where $c$ is the speed of sound. This can be interpreted that a small pressure wave propagates with the speed of sound relative to the flow.

@@

@@

\begin{figure*}[b!]
	%\centering 
	\includegraphics[trim=0cm 6cm 0cm 0cm, clip,height=0.9\textheight]{Sections/ultraReportT5.pdf}
	\caption{The result of solid density measurement}
\end{figure*}

>>>>

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
We consider a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ denotes the state variables, $\Theta$ represents the parameters, and $u$ is the control variables. Our focus is on the supercritical extraction process, which involves a partially filled extractor with a fixed bed that operates under constant conditions. We assume the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. To model this two-dimensional case, we use the concept of quasi-one-dimensional flow.
To describe the fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent, we use a distributed-parameter model based on \citet{Reverchon1996}. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, and saturation concentration. We estimate these parameters from four experiments at different temperatures and pressures: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. We obtain the parameter values using the maximum likelihood estimation method on yield data under the normal error assumption.

@@

\begin{keywords}
Supercritical extraction \sep Parameter estimation \sep Mathematical modeling
\end{keywords}

@@

\subsection{Governing equations}
The governing equation for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}.  Quasi-one-dimensional flow is a fluid flow characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as an irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

@@

Based on governing equations, the small discontinuity (defined as $\delta$) in flow properties, shown in Figure \ref{fig: Discontinuity_slow_flow}, can be analysed. The analysis follows the work of \citet{Schreier1982}.

\begin{figure}[!h]
	\centering
	\resizebox{0.95\columnwidth}{!}{%
		\begin{tikzpicture}[]
			\draw (0,2) -- (9,2);	% Top line
			\draw (0,0) -- (9,0); 	% Bottom line
			\draw (4.5,0) -- (4.5,2); 	% Bottom line
			\node at (4,1.5) {$\rho_f$};
			\node at (5.5,1.5) {$\rho_f+\delta\rho_f$};
			\node at (4,1.0) {$P$};
			\node at (5.5,1.0) {$P+\delta P$};
			\node at (4,0.5) {$T$};
			\node at (5.5,0.5) {$T+\delta T$};
			\node at (1,1.0) {$v \rightarrow$};
			\node at (8,1.0) {$v + \delta v \rightarrow$};
	\end{tikzpicture} }
	\caption{Small discontinuity in one-dimensional flow}
	\label{fig: Discontinuity_slow_flow}
\end{figure}

The discontinuity is presumed to be at rest relative, and the balance equations become

@@

These relations are equally valid if the two regions are separated by a region of finite width rather than a discontinuity.

@@

By combining the momentum equation with the above equation, we get

@@

Suppose the flow is presumed to be isentropic, $dP/d\rho = c^2$, so $v^2=c^2$, where $c$ is the speed of sound. This can be interpreted as a small pressure wave propagating with the speed of sound relative to the flow.

@@

\subsection{Cardano's Formula} \label{CH: Cardano}
\subfile{Sections/Cardano}

\subsection{Porosity Calculations} \label{CH: Porosity}
\subfile{Sections/Porosity}

@@

\begin{figure*}[!h]
	%\centering 
	\includegraphics[trim=0cm 6cm 0cm 0cm, clip,height=0.9\textheight]{Sections/ultraReportT5.pdf}
	\caption{The result of solid density measurement}
\end{figure*}

### 2023-04-22 SFE-article/Sections/Bayes_Theorem.tex
08.03
21.04

\documentclass[../Parameter_fitting.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

As discussed by \citet{Himmelblau1970}, the Bayesian approach to estimation make use of prior information. Such prior knowledge can come from theoretical considerations, from the results of previous experiments, or from assumptions by the experimenter. Typically, a Bayesian approach assumes a prior probability distribution of an unknown parameter $\theta$ in some parameter space $\boldsymbol{\theta}$. The distribution is updated by using Bayes' rule to obtain the posterior probability distribution.

Consider a set of events or outcomes, $A_1,A_2,...,A_n$, and some other event $B$. Bayes' theorem states that the probability that event $A_i$ will occur, given that event $B$ has already occured, which will denoted by $P\{A_i|B\}$, is equal to the product of the probability that $A_i$ will occur regardless of whether $B$ will take place and the probability that $B$ will occur, given that $A_i$  has already taken place, divided by the probability of the occurrence of $B$:

@@

For continuous variable, Bayes' theorem can be more conveniently expressed in terms of the probability density function rather than the probabilities themselves. Equation \ref{EQ: Bayes_discrit} can be expressed in terms of a set of observed values of the random variable $X$, the experimental data \textbf{x} and unknown parameter $\theta$ as

@@

\begin{enumerate}
	\item It follows from the frequency theory of probability that: $0 \leq P \leq 1$
	\item If the probability of occurrence of one event $A$ depends on whether or not event $B$ has occurred, the two events are termed \textit{dependent}; if the probability of occurrence of event $A$ does not depend on the occurrence of $B$, or the reverse, the two events are \textit{independent}.
	\item \textbf{Addition Rule}
	If $A_1,A_2,...,A_n$ are mutually exclusive events, i.e., cannot occur at the same time, the probability of occurrence of just one of the events is equal to the sum of the probabilities of each $A_i$:

@@

	Also, if each event is equi-probable so that $P(A_i) = q$,

@@

	In set theory, mutually exclusive event have no points in common. The union of the sets which represents the set of all elements that belong to

@@

	Two kind of probabilities enter Equation \ref{EQ: Probabiltiy_dependet_a} (or Equation \ref{EQ: Probabiltiy_dependet_b}): the absolute probability of event $B$ (or $A$) irrespective of whether or not $A$ (or $B$) has occurred, and that the conditional probability of event $A$ (or $B$) computed on the assumption that $B$ (or $B$) has occurred. Equation \ref{EQ: Probabiltiy_independet_a} or \ref{EQ: Probabiltiy_independet_b} is special case of Equation \ref{EQ: Probabiltiy_dependet_a} or \ref{EQ: Probabiltiy_dependet_b}, because if the event are independent $P(A|B)=P(A)$.
	For the case of many event, Equation \ref{EQ: Probabiltiy_independet_a} can be expanded to

@@

	\item Another useful relationship for event which are not mutually exclusive is

>>>>

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

As discussed by \citet{Himmelblau1970}, the Bayesian approach to estimation makes use of prior information. Such prior knowledge can come from theoretical considerations, from the results of previous experiments, or from assumptions by the experimenter. Typically, a Bayesian approach assumes a prior probability distribution of an unknown parameter $\theta$ in some parameter space $\boldsymbol{\theta}$. The distribution is updated by using Bayes' rule to obtain the posterior probability distribution.

Consider a set of events or outcomes, $A_1, A_2,..., A_n$, and some other event $B$. Bayes' theorem states that the probability that event $A_i$ will occur, given that event, $B$ has already occurred, which will be denoted by $P\{A_i|B\}$, is equal to the product of the probability that $A_i$ will occur regardless of whether $B$ will take place and the probability that $B$ will occur, given that $A_i$  has already taken place, divided by the probability of the occurrence of $B$:

@@

For a continuous variables, Bayes' theorem can be more conveniently expressed in terms of the probability density function rather than the probabilities themselves. Equation \ref{EQ: Bayes_discrit} can be expressed in terms of a set of observed values of the random variable $X$, the experimental data \textbf{x} and unknown parameter $\theta$ as

@@

\begin{enumerate}
	\item It follows from the frequency theory of probability that: $0 \leq P \leq 1$
	\item If the probability of occurrence of one event $A$ depends on whether or not event $B$ has occurred, the two events are termed \textit{dependent}; if the probability of occurrence of event $A$ does not depend on the occurrence of $B$ or the reverse, the two events are \textit{independent}.
	\item \textbf{Addition Rule}
	If $A_1, A_2,..., A_n$ are mutually exclusive events, i.e., cannot occur at the same time, the probability of occurrence of just one of the events is equal to the sum of the probabilities of each $A_i$:

@@

	Also, if each event is equiprobable so that $P(A_i) = q$,

@@

	In set theory, mutually exclusive events have no points in common. The union of the sets which represents the set of all elements that belong to

@@

	Two kinds of probabilities enter Equation \ref{EQ: Probabiltiy_dependet_a} (or Equation \ref{EQ: Probabiltiy_dependet_b}): the absolute probability of event $B$ (or $A$) irrespective of whether or not $A$ (or $B$) has occurred, and that the conditional probability of event $A$ (or $B$) computed on the assumption that $B$ (or $B$) has occurred. Equation \ref{EQ: Probabiltiy_independet_a} or \ref{EQ: Probabiltiy_independet_b} is special case of Equation \ref{EQ: Probabiltiy_dependet_a} or \ref{EQ: Probabiltiy_dependet_b}, because if the event are independent $P(A|B)=P(A)$.
	For the case of many event, Equation \ref{EQ: Probabiltiy_independet_a} can be expanded to

@@

	\item Another useful relationship for events that are not mutually exclusive is

### 2023-04-22 SFE-article/Sections/Experiments.tex
21.04

	The close the optimization problem given by set of equations \ref{EQ: Optimization_formulation_MLE}, it is require to know the dataset $Y(t)$. In this case, the dataset have been obtained by performing extraction of caraway oil from caraway seeds. The caraway seeds have been obtained from the company Caraway Finland\todo{Should be mentioned?} during 2022 growing season. The caraway seeds have been pre-treated with Retsch SM 300 cutting mill to decrease the particle size to 1mm and break the outer shell of the seeds. Later, the moisture content was analysed to ensure that it does not have a negative effect on the extraction process. An Infrared Moisture Analyser was used to investigate the sample. The moisture content after grinding was 4.83\%. The grind material density was measured by pyncometery and its result can be found in appendix \ref{CH: Solid_Density_Measurment}. Moreover the porosity was calculated to be 0.5, as explained in appendix \ref{CH: Porosity}.

	The unit used for the experimental work is 10 litre extractor \todo{Should Biorukki be mentioned?} with inner diameter around 15 cm and 60 cm height \todo{Double check dimension}. Four experiments have been performed in that unit at following operating conditions: $40^\circ C / 200$ bar, $50^\circ C / 200$ bar, $40^\circ C / 300$ bar and $50^\circ C / 300$ bar. 1 kg (or 1.6 litre) of grind material was weighted and placed in the extraction chamber. Due to technical limitations, the amount of solid material for extraction was selected to be 1 kg per batch, which is not sufficient to fill the whole chamber.

	After the material was loaded into chamber, the extractor is pre-heated to desired temperature. The next step is to close the outlet line coming out of the extractor and fill the extraction chamber with $CO_2$. The $CO_2$ is pumped and compressed until the operating pressure is achieved. When the operating temperature and pressure are obtained, the outlet in is open to allow the solvent flow through the system. The solvent extract essential oils from the fixed bed of solid material and the mixture $CO_2$-oil flow from the extractor to separator. The separator was operating at $50^\circ C$ and $50$ bar. At this condition the $CO_2$ is gasified and leave from the top, while oil stays in the liquid form and it is collected at the bottom of separator. Every 5 minutes, the oil is drained from the separator and its weight is measured. The amount of oil collect during a batch is an output of the experiment and it is used as data for parameter estimation. The extraction time for each batch was 150 minutes, counted from the opening of the extractor outlet port.

	The result of the experiment are presented in the table \ref{tab: Yield_data}.

>>>>

	The close the optimization problem given by a set of equations \ref{EQ: Optimization_formulation_MLE}, it is required to know the dataset $Y(t)$. In this case, the dataset has been obtained by performing the extraction of caraway oil from caraway seeds. The caraway seeds have been obtained from the company Caraway Finland\todo{Should be mentioned?} during the 2022 growing season. The caraway seeds have been pre-treated with Retsch SM 300 cutting mill to decrease the particle size to 1mm and break the outer shell of the seeds. Later, the moisture content was analysed to ensure that it does not have a negative effect on the extraction process. An Infrared Moisture Analyser was used to investigate the sample. The moisture content after grinding was 4.83\%. The grind material density was measured by pycnometer and its result can be found in appendix \ref{CH: Solid_Density_Measurment}. Moreover, the porosity was calculated to be 0.5, as explained in appendix \ref{CH: Porosity}.

	The unit used for the experimental work is a 10-liter extractor \todo{Should Biorukki be mentioned?} with an inner diameter of around 15 cm and 60 cm height \todo{Double check dimension}. Four experiments have been performed in that unit at the following operating conditions: $40^\circ C / 200$ bar, $50^\circ C / 200$ bar, $40^\circ C / 300$ bar and $50^\circ C / 300$ bar. 1 kg (or 1.6 liters) of grind material was weighed and placed in the extraction chamber. Due to technical limitations, the amount of solid material for extraction was selected to be 1 kg per batch, which is not sufficient to fill the whole chamber.

	After the material was loaded into the chamber, the extractor is pre-heated to desired temperature. The next step is to close the outlet line coming out of the extractor and fill the extraction chamber with $CO_2$. The $CO_2$ is pumped and compressed until the operating pressure is achieved. When the operating temperature and pressure are obtained, the outlet in is open to allow the solvent to flow through the system. The solvent extracts essential oils from the fixed bed of solid material and the mixture $CO_2$-oil flow from the extractor to the separator. The separator was operating at $50^\circ C$ and $50$ bar. At this condition, the gasified $CO_2$ leaves from the top, while oil stays in the liquid form and is collected at the bottom of the separator. Every 5 minutes, the oil is drained from the separator, and its weight is measured. The amount of oil collected during a batch is an output of the experiment and it is used as data for parameter estimation. The extraction time for each batch was 150 minutes, counted from the opening of the extractor outlet port.

	The results of experiments are presented in the table \ref{tab: Yield_data}.
